FMA: A Dataset For Music Analysis

### 2017-04-18 fma.tex
proof-reading, final version sent to co-authors

\documentclass{article}
\usepackage[T1]{fontenc} \usepackage[utf8]{inputenc}
\usepackage[svgnames]{xcolor}
\usepackage[bookmarks=false,colorlinks=true,allcolors=DarkBlue]{hyperref} % MidnightBlue, NavyBlue, DarkBlue, Maroon
\usepackage{ismir,amsmath}
\usepackage{graphicx} \graphicspath{{figs/}}
\usepackage{booktabs, threeparttable, tabularx}
\usepackage{enumitem} \setlist{itemsep=0pt,topsep=0pt,parsep=0pt,partopsep=0pt}

@@

% about 150-200 words.
\begin{abstract}
The Free Music Archive (FMA) is an open and easily accessible dataset that can be used to evaluate several tasks in music information retrieval (MIR).
% TODO: which tasks
While content-based MIR tasks have typically been solved with feature engineering,
%with a two-stage approach: engineered features are extracted from music audio signals, and are then used as input to a regressor or classifier.
the community is becoming interested in feature learning and end-to-end learning. %, i.e. learning from signals to labels
That endeavor is however restrained by the limited availability of large audio datasets.
% The primary goal of this dataset is to enable the training of large models, which require lots of training data, which could potentially avert the semantic gap currently observed between the extracted low-level features such as MFCCs and the high-level targets such as genre.
By releasing the FMA we hope to foster research which will improve the state-of-the-art in several tasks and hopefully surpass the performance ceiling observed in e.g. genre recognition (MGR).
% as it happened in computer vision with the release of ImageNet.
The released data is composed of \ntracks tracks, \nartists artists, \nalbums albums, \ngenres genres, for a total of \tduration days of audio and \size GiB, all under permissive Creative Commons licenses.
It features a collection of metadata such as the song title, album, artist name and genres; user data such as play counts and favorites; free-form text such as description, biography and tags; together with full-length and high-quality audio as well as some pre-computed features. We propose a train/validation/test split and three subsets: a small genre-balanced dataset of 8,000 songs from 8 major genres, a medium genre-unbalanced dataset of 25,000 songs from 16 genres, and a 90 GiB version of the full set with clips trimmed to 30s.
This paper describes the dataset and how it was created, proposes some tasks, and evaluates some baselines for MGR.
% Xavier: tasks, such as ...
Code, data, and usage examples are available at \url{https://github.com/mdeff/fma}.
%All the code to create,  as well as links to download the data and usage examples are available at \url{https://github.com/mdeff/fma}.

@@

% which contains 77,643 songs and 68 genres spanning 26.9 days of song listening
\end{abstract}

@@

\begin{table}[t]
	\centering
	\begin{threeparttable}
	\begin{tabular}{l@{ }rrcc}
		\toprule
		dataset\tnote{1} & \#clips & \#artists & year & audio \\
		\midrule
		\href{https://staff.aist.go.jp/m.goto/RWC-MDB/}{RWC} \cite{RWC} & 465 & - & 2001 & yes \\
		\href{http://calab1.ucsd.edu/~datasets/cal500/}{CAL500} \cite{cal500} & 500 & 500 & 2007 & yes \\
		\href{http://mtg.upf.edu/ismir2004/contest/tempoContest/node5.html}{Ballroom} \cite{ballroom} & 698 & - & 2004 & yes \\
		\href{http://ismir2004.ismir.net/genre_contest/}{ISMIR2004} & 729 & - & 2004 & yes \\
		\href{https://marsyasweb.appspot.com/download/data_sets/}{GZTAN} \cite{gtzan} & 1,000 & $\sim300$ & 2002 & yes \\
		\href{http://www.cp.jku.at/datasets/musiclef/}{MusiClef} \cite{musiclef} & 1,355 & 218 & 2012 & yes \\
		\href{https://labrosa.ee.columbia.edu/projects/artistid/}{Artist20} \cite{artist20} & 1,413 & 20 & 2007 & yes \\
		\href{http://www-ai.cs.uni-dortmund.de/audio.html}{Homburg} \cite{homburg} & 1,886 & 1,463 & 2005 & yes \\  % GarageBand
		103-Artists \cite{103artists} & 2,445 & 103 & 2005 & yes \\
		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{Unique} \cite{unique} & 3,115 & 3,115 & 2010 & yes \\
		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{1517-Artists} \cite{1517artists} & 3,180 & 1,517 & 2008 & yes \\
		\href{http://www.ppgia.pucpr.br/~silla/lmd/}{LMD} \cite{lmd} & 3,227 & - & 2007 & no \\
		\href{http://anasynth.ircam.fr/home/media/ExtendedBallroom}{EBallroom} \cite{extballroom} & 4,180 & - & 2016 & no\tnote{2} \\
		\href{https://labrosa.ee.columbia.edu/projects/musicsim/uspop2002.html}{USPOP} \cite{uspop} & 8,752 & 400 & 2003 & no \\
		\href{http://calab1.ucsd.edu/~datasets/cal10k/}{CAL10k} \cite{cal10k} & 10,271 & 4,597 & 2010 & no \\ % Swat10k
		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{TagATune} \cite{magnatagatune} & 25,863\tnote{3} & 230 & 2009 & yes\tnote{4} \\ % MagnaTagATune
		\href{http://jmir.sourceforge.net/index_Codaich.html}{Codaich} \cite{codaich} & 26,420 & 1,941 & 2006 & no \\ % paper 20,849
		\bf \href{https://github.com/mdeff/fma/}{FMA} & \bf \ntracks & \bf \nartists & \bf 2017 & \bf yes \\
		\href{http://www.omras2.org/}{OMRAS2} \cite{omras} & 152,410 & 6,938 & 2009 & no \\
		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} & 1,000,000 & 44,745 & 2011 & no\tnote{2} \\
		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} & 2,084,320 & - & 2017 & no\tnote{2} \\
		\bottomrule
	\end{tabular}
	\begin{tablenotes}
		% 103-Artists is also known as DB-L, Extended Ballroom
		\item[1] Names are links to datasets' homepage.
		\item[2] Audio can be downloaded from \href{http://www.ballroomdancers.com}{ballroomdancers.com}, \href{https://www.7digital.com}{7digital.com}, \href{https://www.youtube.com}{youtube.com}.
		\item[3] The 25,863 clips are cut from 5,405 songs.
		\item[4] Low quality 16 kHz, 32kbps, mono mp3.
	\end{tablenotes}
	\end{threeparttable}
	\caption{Comparison with some other datasets.} % Genre classification datasets.
	\label{tab:datasets}
\end{table}
% Lists of datasets:
% * http://www.audiocontentanalysis.org/data-sets/
% * A Survey of Evaluation in Music Genre Recognition

\begin{table*}[t]
	\centering
	\begin{threeparttable}
		\begin{tabular}{l@{\hspace{2.7em}}r
			@{\hspace{2.7em}}>{$}r<{$}
			@{}>{${}={}$}c@{}
			>{$}l<{$}
			@{\hspace{2.7em}}>{$}c<{$}
			@{\hspace{2.7em}}c@{\hspace{2.7em}}c}
		\toprule
		dataset & \#samples & \multicolumn{3}{c}{dimensionality} & \multicolumn{3}{c}{size} \\
		\cmidrule{6-8}
				&           & \multicolumn{3}{c}{}               & \text{scale} & [GB] & [days] \\
		\midrule
		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{TagATune} \cite{magnatagatune} &
			25,863 & 29\cdot16\cdot10^3\cdot1 && 4.6\cdot10^5 & 1.2\cdot10^{10} & 3 & 8.7 \\ % MagnaTagATune
		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} \tnote{1} &
			2,084,320 & 10\cdot44\cdot10^3\cdot2 && 8.8\cdot10^5 & 1.8\cdot10^{12} & - & 241 \\
		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} \tnote{2} &
			1,000,000 & 49\cdot27\cdot10^3\cdot2 && 2.6\cdot10^6 & 2.6\cdot10^{12} & 625 & 541 \\
		\href{https://github.com/mdeff/fma/}{FMA} \tnote{2} &
			\ntracks & \aduration\cdot44\cdot10^3\cdot2 && 2.4\cdot10^7 & 2.6\cdot10^{12} & \size & \tduration \\
		\midrule
		\href{http://yann.lecun.com/exdb/mnist/}{MNIST} \cite{mnist} &
			70,000 & 28\cdot28\cdot1 && 7.8\cdot10^2 & 5.5\cdot10^{7\phantom0} & 0.011 & - \\
		\href{https://www.cs.toronto.edu/~kriz/cifar.html}{CIFAR-10} \cite{cifar} &
			60,000 & 32\cdot32\cdot3 && 3.0\cdot10^3 & 1.8\cdot10^{8\phantom0} & 0.16 & - \\
		\href{http://www.image-net.org/challenges/LSVRC/2012/}{ILSVRC 2012-17} \cite{imagenet_challenge} \tnote{3,4} &
			1,341,167 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.0\cdot10^{11} & 100 & - \\
		\href{http://www.image-net.org}{ImageNet} \cite{imagenet} \tnote{3} &
			14,197,122 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.5\cdot10^{12} & 1000 & - \\
		\bottomrule
	\end{tabular}
	\begin{tablenotes}
		\item[1] \href{https://support.google.com/youtube/answer/6039860}{YouTube recommended} sample rate for music videos.
		\item[2] Average duration and sample rate. Most excerpts from \href{https://www.7digital.com}{7digital} are 30 or 60s and 22 or 44kHz \cite{msd_features}.
		\item[3] Average resolutions. Most applications resize or crop to $256\cdot256$ for a scale of $2.6\cdot10^{11}$ and $2.8\cdot10^{12}$.
		\item[4] Object detection / classification task. It is the most used dataset for image classification.
	\end{tablenotes}
	\end{threeparttable}
	\caption{Comparison between the largest public audio and image datasets. Dimensionality is (length $\cdot$ sample rate $\cdot$ \#channels) for audio and (xdim $\cdot$ ydim $\cdot$ \#channels) for images. Scale is the number of samples times the dimensionality. Size is for a (zipped) archive of all \texttt{.mp3} or \texttt{.jpg} in GB, an indication of the quantity of information affected by the size and diversity of the data. The last column is the number of days necessary to listen to the whole available audio.}
	\label{tab:size}
\end{table*}
% Magnatagature dead links: http://tagatune.org/Datasets.html http://tagatune.org/Magnatagatune.html
% Announcement: https://musicmachinery.com/2009/04/01/magnatagatune-a-new-research-data-set-for-mir/

@@

% Value proposition.
\textbf{Large scale.} Large datasets are needed to avoid overtraining and to effectively learn models that incorporate the ambiguities and inconsistencies that one finds with genre. They are also more diverse and allows to average out annotation noise.
While FMA features less clips than MSD or AudioSet, every other dataset with readily available quality audio are two orders of magnitude smaller (see \tabref{tab:datasets}).
For statistical learning algorithms, the scale of FMA is comparable to ImageNet and the most used ILSVRC (see \tabref{tab:size}) and shall thus be sufficient for the foreseable future.
Note that while image datasets are larger, the amount of data is similar because of their smaller dimensionality.

\textbf{Permissive licensing.} MIR research has historically suffered from the lack of publicly available benchmark datasets, which stem from the commercial interest in music by record labels, and therefore imposed rigid copyright.
The sole solution to that problem is to aim for songs which license permits redistribution.
Therfore all songs chosen for the FMA are licensed under Creative Commons by the artists. All metadata and code produced by ourselves are licensed under the \href{https://creativecommons.org/licenses/by/4.0)}{
%Creative Commons Attribution 4.0 International License (CC BY 4.0)
CC BY 4.0} and MIT licenses.
% TODO paper CC BY 4.0 if published at ISMIR
% TODO licensing
% Creative Commons: MagnaTagATune
% Proprietary, non-redistributable: MSD, AudioSet?

\textbf{High-quality audio.}
\tabref{tab:datasets} shows that while the smaller datasets are usually distributed with audio, often regardless of copyrights, most of the larger don't.
They either contain features derived from the audio or provide links to download the audio from an online service. The problem with the former is that researchers are stuck with the set of chosen features and prevented to leverage feature learning algorithms or end-to-end learning systems such as DL, while the later  makes no assurance that the files or services won't disappear or change without notice.
Moreover, researchers are wary of proprietary features such as those computed by commercial services like \href{http://the.echonest.com/}{Echonest} and have extracted features by themselves on the MSD by downloading sample audio, a tedious process \cite{msd_features}.
Finally, released or available to download audio is usually clips of 10 to 30 seconds, sometimes of low quality.
In comparison, FMA is distributed with legally redistributable full-length, high-quality and non-modified audio.
% \todo{See Table ? for a quantitative analysis of quality.}
% TODO table quality

@@

The dataset is a dump of the \href{https://freemusicarchive.org/}{Free Music Archive}, an interactive library of high-quality, legal audio downloads directed by \href{https://wfmu.org/}{WFMU}, the longest-running freeform radio station in the United States.
The archive is funded by the New York State Music Fund, the MacArthur Foundation, the National Endowment for the Arts, and from the project's users.
The website provides a large catalog of artists and high-quality mp3-encoded songs, hand-picked by established audio curators. Each song is legally free to download as artists decided to release their works under a permissive license, most tracks being released under a \href{https://creativecommons.org/}{Creative Commons license} which allows redistribution of the audio. Cheyenne Hohman, the director of the FMA, \href{http://freemusicarchive.org/member/cheyenne_h/blog/100000_SONGS}{announced} in July 2016 that the website had reached the landmark of 100,000 songs.
Inspired by Creative Commons and the open-source software movement, the FMA provides a legal and technological platform for curators, artists, and listeners to harness the potential of music sharing. While the archive is free and open to anyone, written and audio content is curated, and permission to upload or edit content is granted on an invitation basis \cite{art:MossFMA}.

@@

The collected track, album and artist meta-informations\footnote{Raw data is available in \texttt{raw\_tracks.csv}, \texttt{raw\_albums.csv}, and \texttt{raw\_artists.csv}.} were cleaned, uniformly formatted, merged and stored in \texttt{tracks.csv}\footnote{See \texttt{creation.ipynb} for the code which created the dataset.} which \tabref{tab:tracks} shows an excerpt. That file is a relational table where each row represents a track and columns are listed in \tabref{tab:metadata}, although some values may be missing for some tracks. The audio for each track is stored in a file wich name is the track id. All songs are mp3-encoded, most of them with sampling rate of 44,100 Hz, bit rate 320 kb/s, and in stereo. Figures~\ref{fig:album_release_year},~\ref{fig:duration_distribution} and \ref{fig:listens_distribution} show the distribution of albums per year, track durations, and play counts per album.
See the \texttt{analysis.ipynb} notebook for a much more detailed analysis of the content.

@@

While it is difficult to devise a hierarchy of genres, we followed that of our source. Indeed, the website offers a parent genre for each of the 163 available genres. \tabref{tab:genres} shows an excerp of that information along with the number of tracks per genre and the associated top-level genre, that is the root of the genre tree. \figref{fig:genre_hierarchy} shows an excerpt of the tree.
In the per-track table, the \texttt{genres} column contain the genre ids indicated by the artist. Then, given such hierarchical information, we constructed a \texttt{genres\_all} column which contains all the genres encountered when traversing the tree from the indicated genres to the roots. Finally, if the traversal yields only one root, the track is given a top-level genre stored in the \texttt{genre\_top} column.
\figref{fig:genre_distribution} shows the distribution of the most common genres. As often encountered with natural categories, it obeys a power law.

@@

The upper part of \tabref{tab:features} lists the (mostly spectral) features we extracted from the audio with the \href{https://github.com/librosa/librosa}{librosa} Python library, version 0.5.0.
Each feature set (except zero-crossing rate) is computed on windows of 2048 samples spaced by hops of 512 samples. Seven statistics were then computed over all windows: the mean, standard deviation, skew, kurtosis, median, minimum and maximum.
Those 518 pre-computed features are distributed in \texttt{features.csv} for all tracks.\footnote{See \texttt{features.py} for the code which computed the features.}

In addition,
% a number of descriptive medium-level and high-level audio
features extracted by \href{http://the.echonest.com/}{Echonest}, a music platform, are provided for some tracks. These features include high-level audio features (acousticness, danceability, energy, instrumentalness, liveness, speechiness, tempo, valence), social features (artist discovery, artist familiarity, artist hotttness, song currency, song hotttness), and MFCC-like temporal features for a number of segments \cite{echonest_temporal}. They are distributed in \texttt{echonest.csv}.
%Echonest audio features are described in Kirell's paper
%Those features can be obtained either by uploading the songs to their website, which are then analyzed and sent back to the user, or simply retrieved from their database if the songs are known to them, i.e. they were analyzed before. All of which can be done through an \href{http://developer.echonest.com}{API}. In our case, we only retrieved features from songs which were already in their database.

@@

With its rich set of metadata, user data, audio and features, the FMA is amenable to many tasks in MIR. We share below some possible uses, which serves to illustrate the breadth of data available in the dataset.

@@

While the original intention of the dataset was to release a large volume of audio for machine learning algorithms, analyzing metadata from such a collection is also extremely interesting.
For instance, one could address questions like: How broads artists are, i.e. to which genres do they contribute to? Do we observe some drift in time for the vocabulary used in song titles, album names or artist names? To which extent tags correspond to genres? What other semantic categories can we found in them?
Other lines of work are to test algorithms for the imputation of missing data, or to study to what extent is the FMA representative of mainstream music, or how similar it is to other datasets.

@@

% Cover song recognition
% Lyrics
% Automatic playlist generation
% style?
% Music classification
	% mood classification,
	% instrument recognition / identification
	% music annotation --> temporal labels, e.g. instrument / note

>>>>

\documentclass{article}
\usepackage[T1]{fontenc} \usepackage[utf8]{inputenc}
\usepackage[svgnames]{xcolor}
\usepackage[bookmarks=false,colorlinks=true,allcolors=DarkBlue]{hyperref} % MidnightBlue, NavyBlue, DarkBlue, Maroon
\usepackage{ismir,amsmath}
\usepackage{graphicx} \graphicspath{{figs/}}
\usepackage{booktabs, threeparttable, tabularx}
\usepackage{enumitem} \setlist{itemsep=0pt,topsep=0pt,parsep=0pt,partopsep=0pt}
\usepackage{scrextend}

@@

% about 150-200 words.
\begin{abstract}
The Free Music Archive (FMA) is an open and easily accessible dataset that can be used to evaluate several tasks in music information retrieval (MIR) such as music classification and annotation or recommendation.
While content-based MIR tasks have typically been solved with feature engineering,
%with a two-stage approach: engineered features are extracted from music audio signals, and are then used as input to a regressor or classifier.
the community is becoming interested in feature learning and end-to-end learning. %, i.e. learning from signals to labels
That endeavor is however restrained by the limited availability of large audio datasets.
% The primary goal of this dataset is to enable the training of large models, which require lots of training data, which could potentially avert the semantic gap currently observed between the extracted low-level features such as MFCCs and the high-level targets such as genre.
By releasing the FMA, we hope to foster research which will improve the state-of-the-art in several tasks and hopefully surpass the performance ceiling observed in e.g.\ genre recognition (MGR).
% as it happened in computer vision with the release of ImageNet.
The released data is composed of \ntracks tracks, \nartists artists, \nalbums albums, \ngenres genres, for a total of \tduration days of audio and \size GiB, all under permissive Creative Commons licenses.
It features a collection of metadata such as the song title, album, artist name and genres; user data such as play counts and favorites; free-form text such as description, biography and tags; together with full-length and high-quality audio as well as some pre-computed features. We propose a train/validation/test split and three subsets: a small genre-balanced dataset of 8,000 songs from 8 major genres, a medium genre-unbalanced dataset of 25,000 songs from 16 genres, and a 90 GiB version of the full set with clips trimmed to 30s.
This paper describes the dataset and how it was created, proposes some tasks, and evaluates some baselines for MGR.
Code, data, and usage examples are available at \url{https://github.com/mdeff/fma}.

@@

\end{abstract}

@@

\begin{table}[t]
	\centering
	\begin{threeparttable}
	\begin{tabular}{l@{ }rrcc}
		\toprule
		dataset\tnote{1} & \#clips & \#artists & year & audio \\
		\midrule
		\href{https://staff.aist.go.jp/m.goto/RWC-MDB/}{RWC} \cite{RWC} & 465 & - & 2001 & yes \\
		\href{http://calab1.ucsd.edu/~datasets/cal500/}{CAL500} \cite{cal500} & 500 & 500 & 2007 & yes \\
		\href{http://mtg.upf.edu/ismir2004/contest/tempoContest/node5.html}{Ballroom} \cite{ballroom} & 698 & - & 2004 & yes \\
		\href{http://ismir2004.ismir.net/genre_contest/}{ISMIR2004} & 729 & - & 2004 & yes \\
		\href{https://marsyasweb.appspot.com/download/data_sets/}{GZTAN} \cite{gtzan} & 1,000 & $\sim300$ & 2002 & yes \\
		\href{http://www.cp.jku.at/datasets/musiclef/}{MusiClef} \cite{musiclef} & 1,355 & 218 & 2012 & yes \\
		\href{https://labrosa.ee.columbia.edu/projects/artistid/}{Artist20} \cite{artist20} & 1,413 & 20 & 2007 & yes \\
		\href{http://www-ai.cs.uni-dortmund.de/audio.html}{Homburg} \cite{homburg} & 1,886 & 1,463 & 2005 & yes \\  % GarageBand
		103-Artists \cite{103artists} & 2,445 & 103 & 2005 & yes \\
		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{Unique} \cite{unique} & 3,115 & 3,115 & 2010 & yes \\
		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{1517-Artists} \cite{1517artists} & 3,180 & 1,517 & 2008 & yes \\
		\href{http://www.ppgia.pucpr.br/~silla/lmd/}{LMD} \cite{lmd} & 3,227 & - & 2007 & no \\
		\href{http://anasynth.ircam.fr/home/media/ExtendedBallroom}{EBallroom} \cite{extballroom} & 4,180 & - & 2016 & no\tnote{2} \\
		\href{https://labrosa.ee.columbia.edu/projects/musicsim/uspop2002.html}{USPOP} \cite{uspop} & 8,752 & 400 & 2003 & no \\
		\href{http://calab1.ucsd.edu/~datasets/cal10k/}{CAL10k} \cite{cal10k} & 10,271 & 4,597 & 2010 & no \\ % Swat10k
		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{TagATune} \cite{magnatagatune} & 25,863\tnote{3} & 230 & 2009 & yes\tnote{4} \\ % MagnaTagATune
		\href{http://jmir.sourceforge.net/index_Codaich.html}{Codaich} \cite{codaich} & 26,420 & 1,941 & 2006 & no \\ % paper 20,849
		\bf \href{https://github.com/mdeff/fma/}{FMA} & \bf \ntracks & \bf \nartists & \bf 2017 & \bf yes \\
		\href{http://www.omras2.org/}{OMRAS2} \cite{omras} & 152,410 & 6,938 & 2009 & no \\
		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} & 1,000,000 & 44,745 & 2011 & no\tnote{2} \\
		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} & 2,084,320 & - & 2017 & no\tnote{2} \\
		\bottomrule
	\end{tabular}
	\begin{tablenotes}
		% 103-Artists is also known as DB-L, Extended Ballroom
		\item[1] Names are clickable links to datasets' homepage.
		\item[2] Audio can be downloaded from \href{http://www.ballroomdancers.com}{ballroomdancers.com}, \href{https://www.7digital.com}{7digital.com}, \href{https://www.youtube.com}{youtube.com}.
		\item[3] The 25,863 clips are cut from 5,405 songs.
		\item[4] Low quality 16 kHz, 32kbps, mono mp3.
	\end{tablenotes}
	\end{threeparttable}
	\caption{Comparison with some other datasets.} % Genre classification datasets.
	\label{tab:datasets}
\end{table}
% Lists of datasets:
% * http://www.audiocontentanalysis.org/data-sets/
% * A Survey of Evaluation in Music Genre Recognition

\begin{table*}[t]
	\centering
	\begin{threeparttable}
		\begin{tabular}{l@{\hspace{2.7em}}r
			@{\hspace{2.7em}}>{$}r<{$}
			@{}>{${}={}$}c@{}
			>{$}l<{$}
			@{\hspace{2.7em}}>{$}c<{$}
			@{\hspace{2.7em}}c@{\hspace{2.7em}}c}
		\toprule
		dataset & \#samples & \multicolumn{3}{c}{dimensionality} & \multicolumn{3}{c}{size} \\
		\cmidrule{6-8}
				&           & \multicolumn{3}{c}{}               & \text{scale} & [GiB] & [days] \\
		\midrule
		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{TagATune} \cite{magnatagatune} &
			25,863 & 29\cdot16\cdot10^3\cdot1 && 4.6\cdot10^5 & 1.2\cdot10^{10} & 3 & 8.7 \\ % MagnaTagATune
		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} \tnote{1} &
			2,084,320 & 10\cdot44\cdot10^3\cdot2 && 8.8\cdot10^5 & 1.8\cdot10^{12} & - & 241 \\
		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} \tnote{2} &
			1,000,000 & 49\cdot27\cdot10^3\cdot2 && 2.6\cdot10^6 & 2.6\cdot10^{12} & 625 & 541 \\
		\href{https://github.com/mdeff/fma/}{FMA} \tnote{2} &
			\ntracks & \aduration\cdot44\cdot10^3\cdot2 && 2.4\cdot10^7 & 2.6\cdot10^{12} & \size & \tduration \\
		\midrule
		\href{http://yann.lecun.com/exdb/mnist/}{MNIST} \cite{mnist} &
			70,000 & 28\cdot28\cdot1 && 7.8\cdot10^2 & 5.5\cdot10^{7\phantom0} & 0.011 & - \\
		\href{https://www.cs.toronto.edu/~kriz/cifar.html}{CIFAR-10} \cite{cifar} &
			60,000 & 32\cdot32\cdot3 && 3.0\cdot10^3 & 1.8\cdot10^{8\phantom0} & 0.16 & - \\
		\href{http://www.image-net.org/challenges/LSVRC/2012/}{ILSVRC 2012-17} \cite{imagenet_challenge} \tnote{3,4} &
			1,341,167 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.0\cdot10^{11} & 100 & - \\
		\href{http://www.image-net.org}{ImageNet} \cite{imagenet} \tnote{3} &
			14,197,122 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.5\cdot10^{12} & 1000 & - \\
		\bottomrule
	\end{tabular}
	\begin{tablenotes}
		\item[1] \href{https://support.google.com/youtube/answer/6039860}{YouTube recommended} sample rate for music videos.
		\item[2] Average duration and sample rate. Most excerpts from \href{https://www.7digital.com}{7digital} are 30 or 60s and 22 or 44kHz \cite{msd_features}.
		\item[3] Average resolutions. Most applications resize or crop to $256\cdot256$ for a scale of $2.6\cdot10^{11}$ and $2.8\cdot10^{12}$.
		\item[4] Object detection / classification task. It is the most used dataset for image classification.
	\end{tablenotes}
	\end{threeparttable}
	\caption{Comparison between the largest public audio and image datasets. Dimensionality is (length $\cdot$ sample rate $\cdot$ \#channels) for audio and (xdim $\cdot$ ydim $\cdot$ \#channels) for images. Scale is the number of samples times the dimensionality. Size is for a (zipped) archive of all \texttt{.mp3} or \texttt{.jpg} in GB, an indication of the quantity of information affected by the size and diversity of the data. The last column is the number of days necessary to listen to the whole available audio.}
	\label{tab:size}
\end{table*}
% Magnatagature dead links: http://tagatune.org/Datasets.html http://tagatune.org/Magnatagatune.html
% Announcement: https://musicmachinery.com/2009/04/01/magnatagatune-a-new-research-data-set-for-mir/

@@

% Value proposition.
\textbf{Large scale.} Large datasets are needed to avoid overtraining and to effectively learn models that incorporate the ambiguities and inconsistencies that one finds with musical categories. They are also more diverse and allows to average out annotation noise.
While FMA features less clips than MSD or AudioSet, every other dataset with readily available quality audio are two orders of magnitude smaller (see \tabref{tab:datasets}).
For statistical learning algorithms, the scale of FMA is comparable to ImageNet and the most used ILSVRC (see \tabref{tab:size}) and shall thus be sufficient for the foreseable future.
Note that while image datasets are larger, the amount of data is similar because of their smaller dimensionality.

\textbf{Permissive licensing.} MIR research has historically suffered from the lack of publicly available benchmark datasets, which stem from the commercial interest in music by record labels, and therefore imposed rigid copyright.
The FMA embraced the sole solution to that problem, that is to aim for tracks which license permits redistribution, such as Creative Commons.
All metadata and code produced by ourselves are licensed under the \href{https://creativecommons.org/licenses/by/4.0)}{
%Creative Commons Attribution 4.0 International License (CC BY 4.0)
CC BY 4.0} and MIT licenses.
% TODO paper CC BY 4.0 if published at ISMIR
% TODO licensing
% Creative Commons: MagnaTagATune
% Proprietary, non-redistributable: MSD, AudioSet?

\textbf{High-quality audio.}
\tabref{tab:datasets} shows that while the smaller datasets are usually distributed with audio, often regardless of copyrights, most of the larger don't.
They either contain features derived from the audio or provide links to download the audio from an online service. The problem with the former is that researchers are stuck with the set of chosen features and are prevented to leverage feature learning algorithms or end-to-end learning systems such as DL, while the later  makes no assurance that the files or services won't disappear or change without notice.
Moreover, researchers are wary of proprietary features such as those computed by commercial services like \href{http://the.echonest.com/}{Echonest} and have extracted features by themselves on the MSD by downloading sample audio, a tedious process \cite{msd_features}.
Finally, released or available to download audio is usually clips of 10 to 30 seconds, sometimes of low quality.
In comparison, FMA is distributed with legally redistributable full-length, high-quality and non-modified audio.
% \todo{See Table ? for a quantitative analysis of quality.}
% TODO table quality

@@

The dataset is a dump of the \href{https://freemusicarchive.org/}{Free Music Archive}, an interactive library of high-quality, legal audio downloads directed by \href{https://wfmu.org/}{WFMU}, the longest-running freeform radio station in the United States.
The archive is funded by the New York State Music Fund, the MacArthur Foundation, the National Endowment for the Arts, and from the project's users.
The website provides a large catalog of artists and high-quality mp3-encoded songs, hand-picked by established audio curators. Each song is legally free to download as artists decided to release their works under a permissive license, most tracks being released under a \href{https://creativecommons.org/}{Creative Commons license} which allows redistribution of the audio. Cheyenne Hohman, the director of the FMA, \href{http://freemusicarchive.org/member/cheyenne_h/blog/100000_SONGS}{announced} in July 2016 that the website had reached the landmark of 100,000 tracks.
Inspired by Creative Commons and the open-source software movement, the FMA provides a legal and technological platform for curators, artists, and listeners to harness the potential of music sharing. While the archive is free and open to anyone, written and audio content is curated, and permission to upload or edit content is granted on an invitation basis \cite{art:MossFMA}.

@@

The collected track, album and artist meta-informations\footnote{Raw data is available in \texttt{raw\_tracks.csv}, \texttt{raw\_albums.csv}, and \texttt{raw\_artists.csv}.} were cleaned, uniformly formatted, merged and stored in \texttt{tracks.csv}\footnote{\label{creation}See \texttt{creation.ipynb} for the code which created the dataset.} which \tabref{tab:tracks} shows an excerpt. That file is a relational table where each row represents a track and columns are listed in \tabref{tab:metadata}, although some values may be missing for some tracks. The audio for each track is stored in a file wich name is the track id. All songs are mp3-encoded, most of them with sampling rate of 44,100 Hz, bit rate 320 kb/s, and in stereo. Figures~\ref{fig:album_release_year},~\ref{fig:duration_distribution} and \ref{fig:listens_distribution} show the distribution of albums per year, track durations, and play counts per album.
See the \texttt{analysis.ipynb} notebook for a much more detailed analysis of the content.

@@

While it is difficult to devise a hierarchy of genres, we followed that of our source. Indeed, the website offers a parent genre for each of the 163 available genres. \tabref{tab:genres} shows an excerp of that information along with the number of tracks per genre and the associated top-level genre, that is the root of the genre tree. \figref{fig:genre_hierarchy} shows an excerpt of the tree.
In the per-track table, the \texttt{genres} column contain the genre ids indicated by the artist. Then, given such hierarchical information, we constructed a \texttt{genres\_all} column which contains all the genres encountered when traversing the tree from the indicated genres to the roots. Finally, if the traversal yields only one root, the track is given a top-level genre stored in the \texttt{genre\_top} column.
\figref{fig:genres_per_track}~and~\ref{fig:genre_distribution} shows the number of genres per track and tracks per genre. As often encountered with natural categories, it obeys a power law.

@@

The upper part of \tabref{tab:features} lists the (mostly spectral) features we extracted from the audio with the \href{https://github.com/librosa/librosa}{librosa} Python library, version 0.5.0 \cite{librosa}.
Each feature set (except zero-crossing rate) is computed on windows of 2048 samples spaced by hops of 512 samples. Seven statistics were then computed over all windows: the mean, standard deviation, skew, kurtosis, median, minimum and maximum.
Those 518 pre-computed features are distributed in \texttt{features.csv} for all tracks.\footnote{See \texttt{features.py} for the code which computed the features.}

In addition,
% a number of descriptive medium-level and high-level audio
features extracted by \href{http://the.echonest.com/}{Echonest} are provided for some tracks. These features (\tabref{tab:features}, lower part) include high-level audio features (acousticness, danceability, energy, instrumentalness, liveness, speechiness, tempo, valence), social features (artist discovery, artist familiarity, artist hotttness, song currency, song hotttness), and MFCC-like temporal features for a number of segments \cite{echonest_temporal}. They are distributed in \texttt{echonest.csv}.
%Echonest audio features are described in Kirell's paper
%Those features can be obtained either by uploading the songs to their website, which are then analyzed and sent back to the user, or simply retrieved from their database if the songs are known to them, i.e. they were analyzed before. All of which can be done through an \href{http://developer.echonest.com}{API}. In our case, we only retrieved features from songs which were already in their database.

@@

With its rich set of metadata, user data, audio and features, the FMA is amenable to many tasks in MIR. We share below some possible uses which serve to illustrate the breadth of data available in the dataset.

@@

While the original intention of the dataset was to release a large volume of audio for machine learning algorithms, analyzing metadata from such a collection is also extremely interesting.
For instance, one could address questions like: How broad artists are, i.e. to which genres do they contribute to? Do we observe some drift in time for the vocabulary used in song titles, album or artist names? To which extent tags correspond to genres? What other semantic categories can we found in them?
Other lines of work are to test algorithms for the imputation of missing data, or to study to what extent is the FMA representative of mainstream music, or how similar it is to other datasets.

@@

% Cover song recognition --> why not, we have track titles
% Lyrics
% Automatic playlist generation
% style?
% Music classification
	% mood classification,
	% instrument recognition / identification
	% music annotation --> temporal labels, e.g. instrument / note
